
% \documentclass[conference]{IEEEtran}
% \documentclass{sig-alternate} % pre 2017
\documentclass[sigconf]{acmart}  % starting in 2017
\input{setup.tex}

\begin{document}

\setcopyright{rightsretained}

%Conference
\copyrightyear{2017}
\acmYear{2017}
\setcopyright{acmlicensed}
% \acmConference{KDD '17}{August 13--17, 2017}{Halifax, NS, Canada}\acmPrice{15.00}\acmDOI{10.1145/3097983.3098195}
\acmConference{KDD'17}{}{August 13--17, 2017, Halifax, NS, Canada.} \acmPrice{15.00} \acmDOI{10.1145/3097983.3098195}
% \acmISBN{978-1-4503-4887-4/17/08}
\acmISBN{ISBN 978-1-4503-4887-4/17/08}

\fancyhead{}  % apparently KDD printer wants no headers so can add their own

% ================================================================
\title{Bolt: Accelerated Data Mining with Fast Vector Compression}
% ================================================================

\author{Davis W. Blalock}
\affiliation{
  \institution{Computer Science and Artificial \\ Intelligence Laboratory}
  \institution{Massachusetts Institute of Technology}
}
\email{dblalock@mit.edu}

\author{John V. Guttag}
\affiliation{
  \institution{Computer Science and Artificial \\ Intelligence Laboratory}
  \institution{Massachusetts Institute of Technology}
}
\email{guttag@mit.edu}

% ------------------------------------------------
\begin{abstract}
% ------------------------------------------------

Vectors of data are at the heart of machine learning and data mining. Recently, vector quantization methods have shown great promise in reducing both the time and space costs of operating on vectors. We introduce a vector quantization algorithm that can compress vectors over $12\times$ faster than existing techniques while also accelerating approximate vector operations such as distance and dot product computations by up to $10\times$. Because it can encode over 2GB of vectors per second, it makes vector quantization cheap enough to employ in many more circumstances. For example, using our technique to compute approximate dot products in a nested loop can multiply matrices faster than a state-of-the-art BLAS implementation, even when our algorithm must first compress the matrices.

In addition to showing the above speedups, we demonstrate that our approach can accelerate nearest neighbor search and maximum inner product search by over $100\times$ compared to floating point operations and up to $10\times$ compared to other vector quantization methods. Our approximate Euclidean distance and dot product computations are not only faster than those of related algorithms with slower encodings, but also faster than Hamming distance computations, which have direct hardware support on the tested platforms. We also assess the errors of our algorithm's approximate distances and dot products, and find that it is competitive with existing, slower vector quantization algorithms. % Finally, we offer theoretical guarantees regarding the performance of our algorithm.

\end{abstract}

% ------------------------------------------------
% CCS taxonomy stuff / keywords
% ------------------------------------------------

\begin{CCSXML}
<ccs2012>
% <concept>
% <concept_id>10002950.10003648</concept_id>
% <concept_desc>Mathematics of computing~Probability and statistics</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
<concept>
<concept_id>10002950.10003648.10003671</concept_id>
<concept_desc>Mathematics of computing~Probabilistic algorithms</concept_desc>
% <concept_significance>300</concept_significance>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002950.10003648.10003688.10003696</concept_id>
<concept_desc>Mathematics of computing~Dimensionality reduction</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10002950.10003705</concept_id>
<concept_desc>Mathematics of computing~Mathematical software</concept_desc>
<concept_significance>100</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

% \ccsdesc[500]{Scalability}
% \ccsdesc[300]{Information Retrieval}
% \ccsdesc[100]{Engineering Challenges}

% \ccsdesc[500]{Mathematics of computing~Probability and statistics}
\ccsdesc[500]{Mathematics of computing~Probabilistic algorithms}
\ccsdesc[300]{Mathematics of computing~Dimensionality reduction}
\ccsdesc[100]{Mathematics of computing~Mathematical software}

\keywords{Vector Quantization, Scalability, Compression, Nearest Neighbor Search}

\maketitle

% ================================================================
\section{Introduction} \label{sec:intro}
% ================================================================

\input{intro.tex}

% % ================================================================
% \section{Definitions and Problem} \label{sec:problem}
% % ================================================================

% \input{problem.tex}

% ================================================================
\section{Related Work} \label{sec:relatedWork}
% ================================================================

\input{related_work.tex}

% ================================================================
% \vspace{-2mm}
\section{Method} \label{sec:method}
% ================================================================

\input{method.tex}

% ================================================================
\section{Experimental Results} \label{sec:results}
% ================================================================

\input{results.tex}

% ================================================================
\vspace{-1mm}
\section{Summary} \label{sec:conclusion}
% ================================================================

We describe Bolt, a vector quantization algorithm that rapidly compresses large collections of vectors and enables fast computation of approximate Euclidean distances and dot products directly on the compressed representations. Bolt both compresses data and computes distances and dot products up to $10\times$ faster than existing algorithms, making it advantageous both in read-heavy and write-heavy scenarios. Its approximate computations can be over $100\times$ faster than the exact computations on the original floating-point numbers, while maintaining correlations with the true values of over $.95$. Moreover, at this level of correlation, Bolt can achieve $10$-$200\times$ compression or more. These attributes make Bolt ideal as a subroutine in algorithms that are amenable to approximate computations, such as nearest neighbor or maximum inner product searches.

It is our hope that Bolt will be used in many production systems to greatly reduce storage and computation costs for large, real-valued datasets.


% ================================================================
\vspace{-1mm}
% References
% ================================================================

% \nocite{*}

% \IEEEtriggeratref{27}	% trigger column break to make cols even
\bibliographystyle{ACM-Reference-Format}
% \bibliographystyle{abbrev}
\bibliography{doc}

\end{document}
